A-maze of Natural Stories: Texts are comprehensible using the Maze task

### 2021-08-11 Analysis/models.Rmd
tweaks

```{r participants}

@@

<!--# Overview

`r data %>% select(subject) %>% unique() %>% nrow()` participants read naturalistic stories from the natural stories corpus. Each participant read 1 story.

We exclude

- participants who do not report English as a native language (`r data_filt %>% select(subject) %>% unique() %>% nrow()` remaining)
- participants who do not get 80% of the words correct (`r data_low_error %>% select(subject) %>% unique() %>% nrow()` remaining)
- practice items (`r data_low_error %>% nrow()` words remaining)
- words that were wrong or were within two after a mistake (`r data_error_free %>% nrow()` words remaining)
- the first word of every sentence (didn't have a real distractor, RT is measured slightly differently) (`r data_no_first %>% nrow()` words remaining)
- words with RTs <100 or >5000 (<100 we think is likely a recording error, or at least not reading the words at all, >5000 is likely getting distracted) (`r data_ready %>% nrow()` words remaining)

Within the filtered data, each story was read between `r min(data_stories$n)` and `r max(data_stories$n)` times, for an average of `r mean(data_stories$n)`.

We also do the analyses on only the words before mistakes (per sentence) (`r data_pre_error %>% nrow()` words)

From the modelling side:
(After attempts without doing this filtering) we only include words which are single token and known words in each of the models vocabularies. We also only include words with frequencies. This is roughly equivalent to excluding words with punctuation.

We use as predictors:

- length in characters of stripped word
- unigram frequency of word. Frequencies for words are calculated using word_tokenize on the gulordava train data and counting up instances. (This tends to tokenize off punctuation, but is capitalization sensitive). Frequencies are represented as log2 of the expected occurances in 1 billion words.

Surprisals are measured in bits.

- ngram (5-gram KN smoothed)
- GRNN
- Transformer-XL

For GAM models, we center length and frequency but not surprisal. We want surprisal interpretable, but we also will be plotting it (at least for the bootstrapping) at length and frequencies set to 0, so they need to be centered. (Not sure this last piece is actually true/matters). -->

# Maze Prep
```{r labels}
labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

@@

```{r maze}

@@

```{r item-means}

@@

```{r for gam}

ngram_data <- select_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                            ) %>% mutate(model="5-gram") 
grnn_data <- select_data %>% select(rt, surprisal=grnn_surp, prev_surp=past_grnn_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center, subject
                                            )%>% mutate(model="GRNN")

txl_data <- select_data %>% select(rt, surprisal=txl_surp, prev_surp=past_txl_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                            ) %>% mutate(model="TXL")

gpt_data <- select_data %>% select(rt, surprisal=gpt_surp, prev_surp=past_gpt_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center,subject
                                            ) %>% mutate(model="GPT-2")

ngram_mean_data <- item_mean_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="5-gram")

@@

all_data <- ngram_data %>% union(grnn_data) %>% union(txl_data) %>% union(gpt_data) %>% 
  select(surprisal, prev_surp,model) %>% 
  pivot_longer(cols=`surprisal`:`prev_surp`) %>% 
  mutate(s=ifelse(name=="surprisal", "Current","Previous"))

@@

```{r formulae, echo=T}
#no hierarchical, has freq x len interaction
formula_interact <- rt ~ s(surprisal, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     ti(freq,len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")+
                     ti(prev_freq, prev_len, bs="cr")

@@

formula_heirarchy <- rt ~ s(surprisal, bs="cr", k=20)+
                    s(surprisal, by=subject, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                      s(prev_surp, by=subject, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")

@@

```{r plots, eval=T}
plot(gam_gpt_1, select=4)
plot(gam_grnn_1, select=4)

plot(gam_txl_1, select=4)
plot(gam_ngram_1, select=4)

@@

It looks like the effect is ~ 0 for most of the space where there's actually data, and it only goes off the rails in the region where there isn't data.

Below is all words with centered frequency < -6 and centered_length > 4. Most of these are I think are recognizable as words. 
```{r}
select_data %>% select(word,  freq_center, length_center, sentence) %>% unique() %>% 
  arrange(freq_center) %>% 
  filter(freq_center< -6 & length_center > 4) %>% knitr::kable()
```

@@

@@

first <- d.raw %>% select(WorkerId,correct,item) %>% unique() %>% 
  group_by(WorkerId,correct) %>% 
  mutate(num=row_number()) %>% 
  filter(num==1) %>% 
  select(-num)

d.limited <- d.raw %>% inner_join(first, by=c("WorkerId","correct","item"))

@@

```{r for spr gam, eval=F}

@@

```{r spr formulae, echo=T}
#no hierarchical, has freq x len interaction
spr_formula_interact <- rt ~ s(surp, bs="cr", k=20)+
                     ti(length_center, bs="cr")+
                     ti(freq_center,length_center, bs="cr")+
                     s(past_surp, bs="cr", k=20)+
                     ti(past_freq_center, bs="cr")+
                     ti(past_length_center, bs="cr")+
                     ti(past_freq_center, past_length_center, bs="cr")+
                     s(past2_surp, bs="cr", k=20)+
                     ti(past2_freq_center, bs="cr")+
                     ti(past2_length_center, bs="cr")+
                     ti(past2_freq_center, past2_length_center, bs="cr")+
                     s(past3_surp, bs="cr", k=20)+
                     ti(past3_freq_center, bs="cr")+
                     ti(past3_length_center, bs="cr")+
                     ti(past3_freq_center, past3_length_center, bs="cr")

@@

```{r spr meaned data, eval=F}
gam_ngram_1 <- gam(spr_formula_interact, data=ngram_mean_spr, method="REML")
gam_grnn_1 <- gam(spr_formula_interact, data=grnn_mean_spr, method="REML")
gam_txl_1 <- gam(spr_formula_interact, data=txl_mean_spr, method="REML")
gam_gpt_1 <- gam(spr_formula_interact, data=gpt_mean_spr, method="REML")

@@

@@

```{r spr brms, eval=F}

@@

#data=readRDS("spr_pre_brms.rds")
data=read_rds(here("Analysis/models/spr_pre_brms.rds"))
 priors <- c(
      set_prior("normal(350,50)", class="Intercept"),
      set_prior("normal(0, 5)", class="b"),
      set_prior("exponential(.01)", class="sd", group="subject", coef="Intercept"),
      set_prior("exponential(.01)", class="sd", group="Word_ID", coef="Intercept"),
     # set_prior("exponential(1)", class="sd"),
     # set_prior("lkj(2)",       class="cor"),
      set_prior("exponential(.01)", class="sigma")
      )
run_model <- function(model_spec,data,file_name){
  d <- data %>% filter(model==model_spec)
  m1 <- brm(rt ~ surp_center * length_center + length_center * freq_center +
                  past_surp_center * past_length_center + past_length_center * past_freq_center + 
                 past2_surp_center * past2_length_center + past2_length_center * past2_freq_center + 
                  past3_surp_center * past3_length_center + past3_length_center * past3_freq_center + 
              (1|subject)+(1|Word_ID),
            data=d,
            file=file_name,
            prior=priors,
            control=list(adapt_delta=.95))
}

#run_model("5-gram",data,"5-gram_spr")
#run_model("GRNN",data,"grnn_spr")
run_model("TXL",data,"txl_spr")
#run_model("GPT-2",data,"gpt_spr")

@@

```
## Summaries of LMER 
These have some heirarchical effects

@@

Really struggling with the intercept for population & subject level -- maybe sd was intercept needs to be a lot bigger?

@@

r_squared <- function(m1){
    lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr) %>% summary()
    lm1$r.squared %>% round(2) %>% as.character()
}

@@

# R-squared

```{r}
spr <- read_rds(here("Analysis/spr_model_compare.rds"))

maze <- read_rds(here("Analysis/maze_model_compare.rds"))
```

@@

Word level correlation between maze and spr
```{r}

summ_spr <- spr %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt))

summ <- maze %>% group_by(Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% left_join(summ_spr)

cor(summ$spr,summ$maze, use="complete.obs")

sentence level correlation between maze and spr

Note that a lot of this is driven purely by sentence length, which isn't interesting. 
```{r}
sentence <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze=sum(maze)/1000,
            spr=sum(spr)/1000)

cor(sentence$spr,sentence$maze, use="complete.obs")
```

Averaging across words at the sentence level removes the sentence-length effect.

```{r}
sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze=mean(maze),
            spr=mean(spr))
ggplot(sentence_avg, aes(x=maze, y=spr))+geom_point(alpha=.1)+
    coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
  labs(title="Sentence level correlation in ms")

cor(sentence_avg$spr,sentence_avg$maze, use="complete.obs")

As a baseline, we can do all the same things on split-halfs of each data set. We split the subjects in half. (Note that this means that each half has the same # of subjects, but may not have the same #s of each story)

Within spr correlations are higher than spr-maze.

```{r spr-split-half}
set.seed(42)
spr_subs <- spr$WorkerId %>% unique()
length(spr_subs)
#180 participants

spr1 <- sample(spr_subs, 90)

spr_split <- spr %>% 
  mutate(half=ifelse(WorkerId %in% spr1, "spr1","spr2")) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=spr)

@@

cor(spr_sentence_avg$spr1,spr_sentence_avg$spr2, use="complete.obs")
```

Same things for maze. Note that with fewer subjects, and given only pre-error data, there's a lot more chance of emptiness. I think the weak sentence average correlation is from places where some sentences didn't go as long in one of the halves (b/c low data).

```{r maze-split-half}
set.seed(42)
maze_subs <- maze$subject %>% unique()
length(maze_subs)
#180 participants

maze1 <- sample(maze_subs, 31)

maze_split <- maze %>% 
  mutate(half=ifelse(subject %in% maze1, "maze1","maze2")) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=maze)

@@

>>>>

# Maze Prep

```{r maze-participants}

@@

```{r labels}
labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

@@

```{r}

@@

```{r maze-item-means}

@@

```{r maze-prep-gam}
#on meaned data
ngram_mean_data <- item_mean_data %>% select(rt, surprisal=ngram_surp, prev_surp=past_ngram_surp, 
                                            freq=freq_center, len=length_center,
                                            prev_freq=past_freq_center, prev_len=past_length_center
                                            ) %>% mutate(model="5-gram")

@@

@@

```{r maze-formulae, echo=T}
#no hierarchical, has freq x len interaction
formula_interact <- rt ~ s(surprisal, bs="cr", k=20)+
                     ti(freq, bs="cr") + 
                     ti(len, bs="cr")+
                     ti(freq,len, bs="cr")+
                     s(prev_surp, bs="cr", k=20)+
                     ti(prev_freq, bs="cr")+
                     ti(prev_len, bs="cr")+
                     ti(prev_freq, prev_len, bs="cr")

@@

@@

## Plots

```{r plots, eval=T}
message("GPT")
plot(gam_gpt_1)
message("GRNN")
plot(gam_grnn_1)
message("TXL")
plot(gam_txl_1)
message("NGRAM")
plot(gam_ngram_1)

@@

For the length: frequency interaction, it looks like the effect is ~ 0 for most of the space where there's actually data, and it only goes off the rails in the region where there isn't data.

<!--Below is all words with centered frequency < -6 and centered_length > 4. Most of these are I think are recognizable as words. -->
```{r}
# select_data %>% select(word,  freq_center, length_center, sentence) %>% unique() %>% 
#   arrange(freq_center) %>% 
#   filter(freq_center< -6 & length_center > 4) %>% knitr::kable()
```

@@

# Maze LM with freq**2 effects

Based on GAM viz, we try models with freq**2 effects on item mean data.

(Tried this with orthagonal polynomials, but I get ridiculous predictions where both the effects for 1st and 2nd order are huge with huge std error, so probably cancelling out?).

Instead, square and then center raw frequency. This also finds the effects, and seems a little absurd, since it's finding big effects, in opposite directions for squared and not squared.

item_mean_data <- select_data %>% 
  group_by(across(c(-subject, -rt))) %>% 
  summarize(rt=mean(rt)) %>% 
  ungroup() %>% 
  mutate(freq_sq_raw=freq**2,
         past_freq_sq_raw=past_freq**2,
         freq_sq=freq_sq_raw-mean(freq_sq_raw, na.rm=T),
         past_freq_sq=past_freq_sq_raw - mean(past_freq_sq_raw, na.rm=T))

no_surp_sq <- "rt ~  freq_sq+ freq_center*length_center + past_freq_sq + past_freq_center*past_length_center"
no_surp_poly <- "rt ~ poly(freq_center,2) + length_center+ freq_center:length_center + poly(past_freq_center,2)+past_length_center + past_freq_center:past_length_center"
ngram <- "+ ngram_center + past_ngram_center + ngram_center:length_center +   
  past_ngram_center:past_length_center"
grnn <- "+ grnn_center + past_grnn_center+ grnn_center:length_center +   
  past_grnn_center:past_length_center"
txl <- "+ txl_center + past_txl_center + txl_center:length_center +   
  past_txl_center:past_length_center"
gpt <- "+ gpt_center + past_gpt_center + gpt_center:length_center +   
  past_gpt_center:past_length_center"

lm(str_c(no_surp_sq,ngram), data=item_mean_data) %>% summary()

lm(str_c(no_surp_sq,grnn), data=item_mean_data) %>% summary()
lm(str_c(no_surp_sq,txl), data=item_mean_data) %>% summary()
lm(str_c(no_surp_sq,gpt), data=item_mean_data) %>% summary()

lm(str_c(no_surp_poly,ngram), data=item_mean_data) %>% summary()
lm(str_c(no_surp_poly,grnn), data=item_mean_data) %>% summary()
lm(str_c(no_surp_poly,txl), data=item_mean_data) %>% summary()
lm(str_c(no_surp_poly,gpt), data=item_mean_data) %>% summary()
```

@@

first <- d.raw %>% select(WorkerId,item) %>% unique() %>% 
  group_by(WorkerId) %>% 
  mutate(num=row_number()) %>% 
  filter(num==1) %>% 
  select(-num)

d.limited <- d.raw %>% inner_join(first, by=c("WorkerId","item"))

@@

```{r for spr gam, eval=T}

@@

```{r spr formulae, echo=T}
#no hierarchical, has freq x len interaction
spr_formula_interact <- rt ~ s(surp, bs="cr", k=20)+
                     ti(length_center, bs="cr")+
                     ti(freq_center,bs="cr")+
                     ti(freq_center,length_center, bs="cr")+
                     s(past_surp, bs="cr", k=20)+
                     ti(past_freq_center, bs="cr")+
                     ti(past_length_center, bs="cr")+
                     ti(past_freq_center, past_length_center, bs="cr")+
                     s(past2_surp, bs="cr", k=20)+
                     ti(past2_freq_center, bs="cr")+
                     ti(past2_length_center, bs="cr")+
                     ti(past2_freq_center, past2_length_center, bs="cr")+
                     s(past3_surp, bs="cr", k=20)+
                     ti(past3_freq_center, bs="cr")+
                     ti(past3_length_center, bs="cr")+
                     ti(past3_freq_center, past3_length_center, bs="cr")

@@

```{r spr meaned data, cache=T, eval=T}
gam_ngram_1 <- gam(spr_formula_interact, data=ngram_mean_spr, method="REML")
gam_grnn_1 <- gam(spr_formula_interact, data=grnn_mean_spr, method="REML")
gam_txl_1 <- gam(spr_formula_interact, data=txl_mean_spr, method="REML")
gam_gpt_1 <- gam(spr_formula_interact, data=gpt_mean_spr, method="REML")

@@

```{r spr-gam-plot}
message("GPT")
plot(gam_gpt_1)
message("GRNN")
plot(gam_grnn_1)
message("TXL")
plot(gam_txl_1)
message("NGRAM")
plot(gam_ngram_1)
```

@@

@@

@@

```
## Summaries of LMER 
These have some hierarchical effects

@@

@@

r_squared <- function(m1){
    lm1 <- lm(str_c(no_surp,m1), data=item_mean_spr) %>% summary()
    lm1$r.squared %>% round(3) %>% as.character()
}

@@

```
# SPR LM with freq**2 effects

no_surp_poly <- "rt ~ poly(freq_center,2) + length_center+ freq_center:length_center + poly(past_freq_center,2)+past_length_center + past_freq_center:past_length_center+
poly(past2_freq_center,2)+past2_length_center + past2_freq_center:past2_length_center+
poly(past3_freq_center,2)+past3_length_center + past3_freq_center:past3_length_center"
ngram <- "+ ngram_center + past_ngram_center + past2_ngram_center + past3_ngram_center+
ngram_center:length_center + past_ngram_center:past_length_center+ past2_ngram_center:past2_length_center+past3_ngram_center:past3_length_center"
grnn <- "+ grnn_center + past_grnn_center +past2_grnn_center +past3_grnn_center +grnn_center:length_center + past_grnn_center:past_length_center+ past2_grnn_center:past2_length_center+past3_grnn_center:past3_length_center"
txl <- "+ txl_center + past_txl_center +past2_txl_center + past3_txl_center +
txl_center:length_center + past_txl_center:past_length_center+ past2_txl_center:past2_length_center+past3_txl_center:past3_length_center"
gpt <- "+ gpt_center + past_gpt_center + past2_gpt_center + + past3_gpt_center+
gpt_center:length_center + past_gpt_center:past_length_center+ past2_gpt_center:past2_length_center+past3_gpt_center:past3_length_center"

lm(str_c(no_surp_poly,ngram), data=item_mean_spr) %>% summary()
lm(str_c(no_surp_poly,grnn), data=item_mean_spr) %>% summary()
lm(str_c(no_surp_poly,txl), data=item_mean_spr) %>% summary()
lm(str_c(no_surp_poly,gpt), data=item_mean_spr) %>% summary()
```

@@

We want to compare w/i Maze, w/i SPR and between Maze and SPR. Maze is the smaller dataset (63 subjects), so we split half for each story.

For SPR, we subsample the 165 subjects down to 63 (to match the number of subjects in Maze), and again split half for each story.

For Maze-SPR comparison, we compare each of the Maze half-SPR half and then average the correlations across these 4 pairs.

### SPR-SPR

```{r spr-split-half}
set.seed(42)
spr_subs <- spr %>% filter(!is.na(Story_Num)) %>% select(WorkerId, Story_Num) %>% unique() %>% sample_n(63) %>%  group_by(Story_Num) %>%
  mutate(half=ifelse(row_number() <= .5*n(),"spr1", "spr2"))
#180 participants

spr_split <- spr %>%
  inner_join(spr_subs) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=spr)

@@

cor(spr_sentence_avg$spr1,spr_sentence_avg$spr2, use="complete.obs")
```
### Maze-Maze

```{r maze-split-half}
set.seed(42)
maze_subs <- maze %>% select(subject, Story_Num) %>% unique() %>% sample_n(63)%>% group_by(Story_Num) %>%
  mutate(half=ifelse(row_number() <= .5*n(),"maze1", "maze2"))

maze_split <- maze %>% 
  left_join(maze_subs) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=maze)

@@

### Maze-SPR
For maze-spr comparisons we take the 4 pairs (maze1 with spr1; maze1 with spr2, etc) We overplot all the combinations (not sure this is right), and take the average of the 4 correlations.

```{r maze-spr}
summ_spr <- spr %>% inner_join(spr_subs)%>% group_by(Story_Num, Sentence_Num, word_num,half) %>% summarize(spr=mean(rt, na.rm=T)) %>% rename(spr_half=half) %>% filter(!is.na(spr))

summ <- maze %>% inner_join(maze_subs) %>% group_by(Story_Num, Sentence_Num, word_num, half) %>% summarize(maze=mean(rt, na.rm=T)) %>% rename(maze_half=half) %>% filter(!is.na(maze))%>% inner_join(summ_spr)

summ11 <- summ %>% filter(maze_half=="maze1" & spr_half=="spr1")
summ12 <- summ %>% filter(maze_half=="maze1" & spr_half=="spr2")
summ21 <- summ %>% filter(maze_half=="maze2" & spr_half=="spr1")
summ22 <- summ %>% filter(maze_half=="maze2" & spr_half=="spr2")

mean(c(cor(summ11$spr,summ11$maze, use="complete.obs"),
  cor(summ12$spr,summ12$maze, use="complete.obs"),
  cor(summ21$spr,summ21$maze, use="complete.obs"),
  cor(summ22$spr,summ22$maze, use="complete.obs")))

sentence <- summ %>% group_by(Story_Num, Sentence_Num, maze_half, spr_half) %>% 
  summarize(maze=sum(maze)/1000,
            spr=sum(spr)/1000)

sentence11 <- sentence %>% filter(maze_half=="maze1" & spr_half=="spr1")
sentence12 <- sentence %>% filter(maze_half=="maze1" & spr_half=="spr2")
sentence21 <- sentence %>% filter(maze_half=="maze2" & spr_half=="spr1")
sentence22 <- sentence %>% filter(maze_half=="maze2" & spr_half=="spr2")

 mean(c(cor(sentence11$spr,sentence11$maze, use="complete.obs"),
  cor(sentence12$spr,sentence12$maze, use="complete.obs"),
  cor(sentence21$spr,sentence21$maze, use="complete.obs"),
  cor(sentence22$spr,sentence22$maze, use="complete.obs")))

sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num, maze_half, spr_half) %>% 
  summarize(maze=mean(maze),
            spr=mean(spr))

sentence_avg11 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr1")
sentence_avg12 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr2")
sentence_avg21 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr1")
sentence_avg22 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr2")

ggplot(sentence_avg, aes(x=maze, y=spr))+geom_point(alpha=.1)+
    coord_cartesian(xlim=c(0,1500),ylim=c(0,600))+
  labs(title="Sentence level correlation in ms")

  mean(c(cor(sentence_avg11$spr,sentence_avg11$maze, use="complete.obs"),
  cor(sentence_avg12$spr,sentence_avg12$maze, use="complete.obs"),
  cor(sentence_avg21$spr,sentence_avg21$maze, use="complete.obs"),
  cor(sentence_avg22$spr,sentence_avg22$maze, use="complete.obs")))
```

### 2021-08-11 Papers/Paper/manuscript.Rmd
tweaks

@@

@@

Many participants completed the Maze task with a high degree of accuracy and also answers the comprehension questions correctly.

@@

Another check is whether participants comprehended the story. We counted how many of the binary-choice comprehension questions each participant got right (out of 6). As seen in Figure \@ref(fig:errors)B, most participants were were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions. Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet. We can't rule out that some participants may have been able to guess the answers without understanding the story. Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task. The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read. Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads. Note that these differences cannot be directly attributed to methods, as the participant populations differed. While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected.

@@

knitr::kable(summ, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models. All terms were centered, but not rescaled. Units are in ms. Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words.")
```

@@

```
# Discussion

>>>>

As described in the Results section, we base all models only on data from participants who choose correctly on at least 80% of the words in the Maze task.

@@

## SPR-Maze correlation
We additionally compare how correlated the Maze and SPR results are to each other, in comparison to within-Maze and within-SPR correlations. For Maze, within each story, we randomly split subjects into two halves. Within each half, we calculate a per-word average for each word and then a per-sentence average RT across word averages. We calculate a within-Maze correlation between these two halves. To avoid differences due to dataset size, we downsample the SPR data choosing a number of participants equal to the number we have for Maze. We then use the same procedure on this subset as for Maze to get a within-SPR correlation. For between Maze-SPR correlation, we take the average correlation across each of the 4 pairs of Maze half and SPR half.

@@

Many participants completed the Maze task with a high degree of accuracy and also answered the comprehension questions correctly.

@@

Another check is whether participants comprehended the story. We counted how many of the binary-choice comprehension questions each participant got right (out of 6). As seen in Figure \@ref(fig:errors)B, most participants who were accurate on the task also did well on comprehension questions, while participants who were at chance on the Maze task were also at chance on the comprehension questions. Participants usually answered quickly (within 10 seconds), so we do not believe they were looking up the answers on the internet. We can't rule out that some participants may have been able to guess the answers without understanding the story. Nonetheless, this provides preliminary evidence that people can understand and remember details of stories they read during the Maze task. The comprehension question performance of accurate Maze participants is broadly similar to the performance of SPR participants from @futrellNaturalStoriesCorpus2020 on the first story read. Overall, `r comp_tally %>% filter(source=="all Maze") %>% pull(good)`% of Maze participants got 5 or 6 questions right (`r comp_tally %>% filter(source=="<80% correct") %>% pull(good)`% of low-accuracy participants and `r comp_tally %>% filter(source==">80% correct") %>% pull(good)`% of high-accuracy participants) compared to `r comp_tally %>% filter(source=="SPR") %>% pull(good)`% of all SPR reads and `r comp_tally %>% filter(source=="SPR 1st story") %>% pull(good)`% of 1st SPR reads. Note that these differences cannot be directly attributed to methods, as the participant populations differed. While both studies were conducted on Mturk, the quality of Mturk data has decreased from 2011 when the SPR was collected to 2020 when the A-maze was collected.

@@

knitr::kable(summ, format="latex", position="ht",caption="Predictions from fitted Bayesian regression models. All terms were centered, but not rescaled. Units are in ms. Surprisal is per bit, length per character, and frequency per $log_2$ occurance per billion words. Interval is 2.5th quantile to 97.5th quantile of model draws.")
```

@@

```{r}
tokenization <- here("Data/SPR/all_stories.tok.txt") %>% 
  read_delim(delim="\t") %>% 
  mutate(Word_In_Story_Num=zone,
         Story_Num=item)

labs <- read_rds(here("Prep_code/natural_stories_surprisals.rds")) %>% 
  left_join(read_delim(here("Materials/natural_stories_sentences.tsv"), delim="\t")) %>% 
  select(word_num=Word_In_Sentence_Num, word=Word, sentence=Sentence, everything())

spr <- read_rds(here("Data/SPR/first.rds")) %>% left_join(tokenization) %>% left_join(labs) %>% select(WorkerId,rt,Story_Num,Sentence_Num, word_num)

maze <- read_rds(here("Data/maze_pre_error.rds")) %>% left_join(labs) %>% select(subject, rt, Story_Num, Sentence_Num, word_num)
```

```{r spr-split-half}
set.seed(42)
spr_subs <- spr %>% filter(!is.na(Story_Num)) %>% select(WorkerId, Story_Num) %>% unique() %>% sample_n(63) %>%  group_by(Story_Num) %>%
  mutate(half=ifelse(row_number() <= .5*n(),"spr1", "spr2"))
#180 participants

spr_split <- spr %>%
  inner_join(spr_subs) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(spr=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=spr)

spr_sentence_avg <- spr_split%>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(spr1=mean(spr1),
            spr2=mean(spr2))

spr_spr <- cor(spr_sentence_avg$spr1,spr_sentence_avg$spr2, use="complete.obs")
```

```{r maze-split-half}
set.seed(42)
maze_subs <- maze %>% select(subject, Story_Num) %>% unique() %>% sample_n(63)%>% group_by(Story_Num) %>%
  mutate(half=ifelse(row_number() <= .5*n(),"maze1", "maze2"))

maze_split <- maze %>% 
  left_join(maze_subs) %>% 
  group_by(half,Story_Num, Sentence_Num, word_num) %>% summarize(maze=mean(rt)) %>% 
  pivot_wider(names_from=half, values_from=maze)

maze_sentence_avg <- maze_split%>% group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze1=mean(maze1),
            maze2=mean(maze2))

maze_maze <- cor(maze_sentence_avg$maze1,maze_sentence_avg$maze2, use="complete.obs")
```

```{r maze-spr}
summ_spr <- spr %>% inner_join(spr_subs)%>% group_by(Story_Num, Sentence_Num, word_num,half) %>% summarize(spr=mean(rt, na.rm=T)) %>% rename(spr_half=half) %>% filter(!is.na(spr))

summ <- maze %>% inner_join(maze_subs) %>% group_by(Story_Num, Sentence_Num, word_num, half) %>% summarize(maze=mean(rt, na.rm=T)) %>% rename(maze_half=half) %>% filter(!is.na(maze))%>% inner_join(summ_spr)

sentence_avg <- summ %>% group_by(Story_Num, Sentence_Num, maze_half, spr_half) %>% 
  summarize(maze=mean(maze),
            spr=mean(spr))

sentence_avg11 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr1")
sentence_avg12 <- sentence_avg %>% filter(maze_half=="maze1" & spr_half=="spr2")
sentence_avg21 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr1")
sentence_avg22 <- sentence_avg %>% filter(maze_half=="maze2" & spr_half=="spr2")

spr_maze <-   mean(c(cor(sentence_avg11$spr,sentence_avg11$maze, use="complete.obs"),
  cor(sentence_avg12$spr,sentence_avg12$maze, use="complete.obs"),
  cor(sentence_avg21$spr,sentence_avg21$maze, use="complete.obs"),
  cor(sentence_avg22$spr,sentence_avg22$maze, use="complete.obs")))
```

(ref:scatter) Correlation between SPR and Maze data. RTs were averaged across participants per word and then averaged together within each sentence. RTs in ms.

```{r scatter, out.height="20%", fig.width=6, fig.height=2, fig.pos="ht", fig.cap="(ref:scatter)"}
spr_collapse <- spr %>% 
  group_by(Story_Num, Sentence_Num,word_num) %>% 
  summarize(spr=mean(rt, na.rm=T)) %>% 
  group_by(Story_Num, Sentence_Num) %>%
  summarize(spr=mean(spr))

both <- maze %>% 
  group_by(Story_Num, Sentence_Num, word_num) %>% 
  summarize(maze=mean(rt, na.rm=T)) %>% 
  group_by(Story_Num, Sentence_Num) %>% 
  summarize(maze=mean(maze)) %>% 
  inner_join(spr_collapse)
ggplot(both, aes(x=maze, y=spr))+geom_point(alpha=.5, size=1)+labs(x="A-Maze", y="SPR")+coord_fixed(ratio=1)
```

An additional comparison between SPR and Maze is how correlated their reading times are; that is, do the methods pick up on the same effects. If so, we would expect that the sentences that take longer to read in one method also take longer in the other. We calculate the average RT at the sentence level (see Methods for details). The correlation between Maze and SPR is `r round(spr_maze,2)`, compared to `r round(spr_spr,2)` within SPR and `r round(maze_maze,2)` within Maze. See Figure \@ref(fig:scatter) for a visual comparison of overall Maze versus SPR RTs.

# Discussion
